---
title: "Data Analysis 2 - Assignment 2"
author: "Tunay Tokmak, Akos Almasi"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Let's clean the environment and import the required libraries.

```{r}
rm(list = ls())
library(tidyverse)
library(modelsummary)
library(marginaleffects)
```

Import the required data sets, filter for Barcelona, and merge them.

```{r}
hotels <- read_csv('hotels-europe_features.csv')
prices <- read_csv('hotels-europe_price.csv')
hotels_prices <- inner_join(hotels,prices, by = 'hotel_id')
hotels_barcelona <- hotels_prices %>% filter(city == 'Barcelona')
```

Normalize the price column. The price for each hotel must present the price per night for a correct analysis.

```{r}
hotels_barcelona <- hotels_barcelona %>% mutate(price_per_night = price / nnights)
```


Evaluate hotel ratings that have more than 5 commentors.

```{r}
hotels_barcelona <- hotels_barcelona %>% filter(rating_reviewcount >= 5)
```

Evaluate hotels less than 5 miles far from center 

```{r}
hotels_barcelona <- hotels_barcelona %>% filter(distance <= 5)
```
Introduce the dummy variable
```{r}
hotels_barcelona <-  hotels_barcelona %>% mutate(highly_rated = ifelse(rating >= 4,1,0))
```

Split the data set into training and testing data sets. Our aim is prediction. Therefore, we split the data using 1/4 ratio.

```{r}
sample <- sample(c(TRUE, FALSE), nrow(hotels_barcelona), replace=TRUE, prob=c(0.8,0.2))
train  <- hotels_barcelona[sample, ]
test   <- hotels_barcelona[!sample, ]
```
